---
title: "Week 3 homework"
author: "Gigi Melone"
date: "9/22/2022"
output: github_document
---

## Loading data and packages 
```{r setup}
library(dplyr)
#install.packages("nycflights13")
library(nycflights13)
library(ggplot2)
```

Question 1: How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
#missing dep time will be NA, so I will look for NA (is.na = TRUE), and then use n() to find the number of missing departure times. 

flights %>% 
  mutate(missing_dep_time = is.na(dep_time)) %>% 
  filter(missing_dep_time == TRUE) %>% 
  summarise(missing_dep_time = n())

```
Answer: 8255 flights are missing dep_time. When dep_time is missing, so are other variables corresponding to flight departure/arrival stats (including departure delay, arrival time, arrival delay, and air time). Variables for scheduled arrival/departure time and flight distance are not missing. I believe the missing rows are present because the flight was cancelled. 

Question 2: Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
#since midnight: 2400 is 1440, add 100 %% 1440 because that's 0 
flights %>% 
  mutate(dep_time_conv = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440, sched_dep_time_conv = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440) %>%
  select(sched_dep_time, sched_dep_time_conv, dep_time, dep_time_conv)
```

Question 3: Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()
```{r}
#H1: greater delays equate to greater cancellations 
delayed_flights_trend <- flights %>% 
  mutate(cancelled = is.na(dep_time) & is.na(arr_time)) %>% 
  group_by(year, month, day) %>% 
  summarise(prop_cancelled = mean(cancelled), avg_dep_delay = mean(dep_delay, na.rm = T), avg_arr_delay = mean(arr_delay, na.rm = T)) 

#plot average delays by proportion cancelled
ggplot(data = delayed_flights_trend, aes(x = avg_dep_delay, y = prop_cancelled)) + geom_point() + geom_smooth()

#accept H1, there is a positive realtionship with length of departure delay and proportion of flights cancelled 
```
Answer: The proportion of cancelled flights is related to average delay, with the proportion of cancelled flights increasing as departure delay increases